Parsing Postfix log files

### 2009-04-08 doc/logparser-results.tex
Worked on caching compiled regexes, and I'm pretty happy with it. Referred to the strange performance parsing the first log file in other sections.

\Graphref{Parsing time relative to shuffled ordering graph} shows the
parsing times of optimal and reverse orderings relative to shuffled
ordering; the mean relative parsing times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  This optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
Optimal rule ordering has other benefits, described in
\sectionref{scalability as the number of rules rises}.  \Tableref{Parsing
time relative to shuffled ordering table} shows that differences in rule
ordering have less effect on parsing time when parsing log files 22 \&
62--68, due to the different distribution of log lines in those log files.

@@

Once again, the difference between the perfect best oracle and optimal
ordering is at its lowest when parsing log files resulting from a mail loop
(log files 22 \& 62--68), because the average number of rules tried per log
line is lower (see \graphref{Average number of rules tried per log line}).
The performance of optimal ordering when parsing the first log file is much
worse than the remainder of the log files, because the hits field of every
rule is zero, so optimal ordering performs similarly to shuffled ordering
for that log file.

@@

Clearly the increased number of rules has a noticeable performance impact
with reverse ordering, and a lesser impact with shuffled ordering, whereas
optimal ordering shows very little change.  Log files 22 \& 62--68 show
much smaller increases than other log files do, because the majority of log
lines in those log files are produced by Postfix components with a small
number of rules, so removing unnecessary rules has little effect on the
total number of rules used; \tableref{Number of rules per Postfix component
in the maximum and minimum rulesets} shows the number of rules per Postfix
component for each ruleset.  The optimal ordering shows a mean increase of
\input{build/include-full-ruleset-vs-minimum-ruleset-mean} in parsing time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that the architecture scales very well as the number of
rules increases, and that optimally sorting the rules is an important
optimisation contributing to this scalability.

@@

Before the Perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and optimised
to improve the speed of matching.  This compilation and optimisation takes
CPU time: in many cases it takes far more CPU time than the actual
matching.  If the interpreter is certain that a regex will not change, it
will automatically cache the compilation results for later use.   The
results of compiling a dynamically generated regex can be cached and used
in preference to the original regex, but it is the responsibility of the
programmer to do this.  \parsername{} loads the rules it uses from a
database, so it compiles each rule's regex and caches the result to reduce
parsing time.

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the impact that not caching compiled regexes has on parser
performance: on typical log files, the parsing time when not caching
compiled regexes is 400--600\% of the parsing time when caching.  Caching
compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement.

@@

XXX UPDATE THIS\@.  Two large dips can be seen in \graphref{Increase in
parsing time when not caching compiled regexes graph} at log files 22 \&
62--68, corresponding to the spikes in log file size in \graphref{parsing
time vs file size vs number of log lines graph}.  The distribution of log
lines across rules when there is a mail loop is much different to the norm,
and the average number of rules consulted per log line is much lower: this
results in far fewer regex compilations per line when there is a mail loop,
and a corresponding decrease in the average parsing time for a single log
line.  The outcome is that caching compiled regexes is proportionally less
important when the log file contents were created by a mail loop.  The
increases in parsing time when not caching compiled regexes for different
groups of log files are summarised in \tableref{Increase in parsing time
when not caching compiled regexes table}.

>>>>

\Graphref{Parsing time relative to shuffled ordering graph} shows the
parsing times of optimal and reverse orderings relative to shuffled
ordering; the mean relative parsing times for different groupings of log
files are given in \tableref{Parsing time relative to shuffled ordering
table}.  This optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files,
\input{build/include-optimal-ordering-parsing-time-reduction-logs-22-62-68}\%
when a mail loop occurs and the distribution of log lines is unusual.
\Tableref{Parsing time relative to shuffled ordering table} shows that
differences in rule ordering have less effect on parsing time when parsing
log files 22 \& 62--68, due to the different distribution of log lines in
those log files.  A careful examination of \graphref{Parsing time relative
to shuffled ordering graph} shows that, for the first log file, optimal and
reverse orderings perform identically: this is because the hits field of
each rule is zero for the first log file, so optimal and reverse orderings
are identical.  For the first log file, shuffled ordering is most
efficient, but that is accidental and cannot be relied upon.

@@

Once again, the difference between the perfect best oracle and optimal
ordering is at its lowest when parsing log files resulting from a mail loop
(log files 22 \& 62--68), because the average number of rules tried per log
line is lower (see \graphref{Average number of rules tried per log line}).
The performance of optimal ordering when parsing the first log file is much
worse than the remainder of the log files, because the hits field of every
rule is zero, so optimal ordering does not provide any benefit for that log
file.

@@

Clearly the increased number of rules has a noticeable performance impact
with reverse ordering, and a lesser impact with shuffled ordering, whereas
optimal ordering shows very little change.  Log files 22 \& 62--68 show
much smaller increases than other log files do, because the majority of log
lines in those log files are produced by Postfix components with a small
number of rules, so removing unnecessary rules has little effect on the
total number of rules used; \tableref{Number of rules per Postfix component
in the maximum and minimum rulesets} shows the number of rules per Postfix
component for each ruleset.  Once again, for the first log file, optimal
and reverse orderings have identical performance, because the hits field of
every rule is zero.

The optimal ordering shows a mean increase of just
\input{build/include-full-ruleset-vs-minimum-ruleset-mean} in parsing time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that the architecture scales very well as the number of
rules increases, and that optimally sorting the rules is an important
optimisation contributing to this scalability.

@@

Before the Perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and optimised
to improve the speed of matching.  This compilation and optimisation takes
CPU time: in many cases it takes far more CPU time than the actual
matching.  If the interpreter is certain that a regex will not change while
the program is running, it will automatically cache the results of
compiling and optimising the regex for later use.  The results of compiling
a dynamically generated regex can be cached and used in preference to the
original regex, but it is the responsibility of the programmer to do this;
\parsername{} does this with every rule's regex when the rules are loaded
from the database.

\Graphref{Increase in parsing time when not caching compiled regexes graph}
shows the impact that not caching compiled regexes has on parser
performance, with mean and standard deviation in \tableref{Increase in
parsing time when not caching compiled regexes table}.  For typical log
files, the mean increase in parsing time when not caching compiled regexes
is \input{build/include-cached-regexes-vs-discarded-regexes-mean}.  Caching
compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement.  As with
previous optimisations, log files 22 \& 62--68 do not suffer such a large
increase in parsing time when not caching compiled regexes; this is
because, on average, fewer regexes are compiled per log line for those log
files.  The increase in parsing time when parsing the first log file is
much greater than for the other log files; again, this is because every
rule's hits field is zero, so optimal ordering is less efficient than
usual, and the average number of rules tried for each log line will be
higher than usual.

@@
